---
title: "Textmining and wordclouds"
output: html_document
date: "2025-05-13"
subtitle: "Analysing data from Mediestream and creating wordclouds"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Organising the workplace
### Creating folders
```{r dir.create, warning=FALSE}
dir.create("data")
dir.create("figures")
```

### Loading libraries
```{r libraries, warning=FALSE}
library(tidyverse)
library(tidytext)
library(ggwordcloud)
library(urltools)
```

### Getting data into R
#### Havesting data directly from data folder
```{r loading dataset 1, warning=FALSE}
kolera_words <- read_csv("data/mediestream_aviser.csv")

```



# Analysing data from Mediestream

### Analysing the publishing place and glimpsing the data
```{r publishing place and glimpse, warning=FALSE}
kolera_words %>%
  count(lplace, sort = TRUE)

glimpse(kolera_words)

```

### Sorting by publisher / publishing house
```{r publishing house, warning=FALSE}
kolera_words %>% 
  count(familyId, sort = TRUE)
```



# Text mining

### Creating a second dataframe called kolera_tidy 
#### Each word now have their own seperate column
```{r creating tidy dataframe, warning=FALSE}
kolera_tidy <- kolera_words %>% 
  unnest_tokens(word, fulltext_org)
```

### Counting the most frequent words without a stopwordlist
```{r count without a stopwordlist, echo=TRUE, warning=FALSE}
kolera_tidy %>% 
  count(word, sort = TRUE)
```

### Loading stopwordlist
```{r stopwordlist, echo=TRUE, warning=FALSE}
stopord_kolera <- read_csv("data/stopwordlist.csv")
```

### Counting the most frequent words, including stopwordlist
```{r count and stopwordlist, echo=TRUE, warning=FALSE}
kolera_tidy %>%
  anti_join(stopord_kolera, by ="word") %>%
  count(word, sort = TRUE)
```






# Wordcloud 1 - most frequent words by year, 1850-1859

### Creating a new dataframe called kolera_year_tidy
#### Contains a new column, with years as values  
```{r kolera_year_tidy, warning=FALSE}
kolera_tidy %>% 
  mutate(y=year(timestamp)) -> kolera_year_tidy
```


### Counting the most frequent words per year, with stopwordlist
```{r most frequent words per year, warning=FALSE}
kolera_year_tidy %>% 
  anti_join(stopord_kolera, by ="word") %>%
  count(word, y, sort = TRUE)
```


### Creating new dataframe called total_words
#### Contains two columns; total mentitioning of words per year and year
```{r total_words, warning=FALSE}
kolera_year_tidy %>% 
  count(word, y) %>% 
  group_by(y) %>% 
  summarise(total=sum(n)) -> total_words

```


### Creating new dataframe calld kolera_year_count
#### Contains four columns; word, year, n-value and total mentioning of the word
#### The n-value is how many times the word was mentioned in a given year
```{r kolera_year_count, warning=FALSE}
kolera_year_tidy %>% 
  count(word, y, sort = TRUE) %>% 
  left_join(total_words, by = "y") -> kolera_year_count

```


### Creating new dataframe called kolera_year_if_idf
#### Couting the mentioning of the acccording to year, including stopwordlist
#### Creating a frequency of the mentioned words per year, compared to the total orrence of words
```{r kolera_year_if_idf, warning=FALSE}
kolera_year_count %>% 
  anti_join(stopord_kolera, by ="word") %>%
  bind_tf_idf(word, y, n) -> kolera_year_if_idf

```


### Arranging the words in descending order by frequency value
```{r descending order, warning=FALSE}
kolera_year_if_idf %>% 
  arrange(desc(tf_idf))

```


### Creating a wordcloud with the most frequent words sorted by year
#### the figure is saved in the figures folder
```{r wordcloud 1, warning=FALSE}
kolera_year_if_idf %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(y) %>% 
  top_n(8) %>% 
  ungroup %>%
  ggplot(aes(label = word, size = tf_idf, color = tf_idf)) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 10) +
  theme_minimal() +
  facet_wrap(~y, ncol = 4, scales = "free") +
  scale_color_gradient(low = "hotpink", high = "black") +
  labs(
      title = "Most frequent words per year from 1850-1859",
       subtitle = "Determined by term frequency (tf) and inversed document frequency(idf)",
      caption = "Data from Mediestream Experimental API"
      )

ggsave("figures/Wordcloud_1.png", width = 18, height = 13, units = "cm", bg = "white")
```







# Wordcloud 2 - Most frequent words by month in the year 1853
### Creating a new dataframe called kolera_1853_tidy
#### Including words only from the year 1853
```{r kolera_1853_tidy, warning=FALSE}
kolera_tidy %>% 
  filter(year(timestamp) == 1853) -> kolera_1853_tidy

```

### Creating a new dataframe called kolera_month_1853_tidy
#### Creating a new column containing months in 1853
```{r kolera_month_1853_tidy, warning=FALSE}
kolera_1853_tidy %>% 
  mutate(m=month(timestamp)) -> kolera_month_1853_tidy

```

### Counting words per month in 1853 without stopwordlist
```{r count - words per month - 1853, warning=FALSE}
kolera_month_1853_tidy %>% 
  count(word, m, sort = TRUE)

```


### Creating a new dataframe called total_words_1853
#### Couting total words from 1853 grouped by month
```{r total_words_1853, warning=FALSE}
kolera_month_1853_tidy %>% 
  count(word, m) %>% 
  group_by(m) %>% 
  summarise(total=sum(n)) -> total_words_1853

```


### Creating a new dataframe called kolera_month_1853_count
#### Counting the words from 1853, by month, without a stopwordlist
```{r kolera_month_1853_count, warning=FALSE}
kolera_month_1853_tidy %>% 
  count(word, m, sort = TRUE) %>% 
  left_join(total_words_1853, by = "m") -> kolera_month_1853_count

```


### Creating a new framework called kolera_month_1853_if_idf
#### Creating a frequency of the most mentioned words per month compared ro the total mentioning in 1853
```{r kolera_month_1853_if_idf, warning=FALSE}
kolera_month_1853_count %>% 
  anti_join(stopord_kolera, by ="word") %>%
  bind_tf_idf(word, m, n) -> kolera_month_1853_if_idf

```


### Arranging the words in descending order by frequency value
```{r descending order - 1853, warning=FALSE}
kolera_month_1853_if_idf %>% 
  arrange(desc(tf_idf))

```


### Creating a wordcloud and visulizing the most mentioned words per word in 1853, including stopwordlist
#### The figure is saved in the figures folder
```{r wordcloud 2, warning=FALSE}
kolera_month_1853_if_idf %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(m) %>% 
  top_n(8) %>% 
  ungroup %>%
  ggplot(aes(label = word, size = tf_idf, color = tf_idf)) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 15) +
  theme_minimal() +
  facet_wrap(~m, ncol = 4, scales = "free") +
  scale_color_gradient(low = "blue", high = "darkblue") +
  labs(
      title = "Most frequent words per month in 1853",
       subtitle = "Determined by term frequency (tf) and inversed document frequency(idf)",
      caption = "Data from Mediestream Experimental API")

ggsave("figures/Wordcloud_2.png", width = 18, height = 13, units = "cm", bg = "white")
```


# Wordcloud 3 - most frequent words in July, August and September in 1853
### Creating a snapshot of the months July, August and September in 1853
#### The figure is saved in the figures folder
```{r wordcloud 3, warning=FALSE}
kolera_month_1853_if_idf %>%
  filter((m %in% 7:9))%>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(m) %>% 
  top_n(8) %>% 
  ungroup %>%
  ggplot(aes(label = word, size = tf_idf, color = tf_idf)) +
  geom_text_wordcloud_area() +
  scale_size_area(max_size = 17) +
  theme_minimal() +
  facet_wrap(~m, ncol = 4, scales = "free") +
  scale_color_gradient(low = "blue", high = "darkblue") +
  labs(
      title = "Most frequent words in July, August and September in 1853",
       subtitle = "Determined by term frequency (tf) and inversed document frequency(idf)",
      caption = "Data from Mediestream Experimental API")

ggsave("figures/Wordcloud_3.png", width = 18, height = 13, units = "cm", bg = "white")
```



# Graph 1 - Occurence of any word in relation to cholera in 1853

### Creating a new dataframe called kolera_1853_mutated_words
#### Contains all words with any connection to cholera, merged into one
#### Created with data from the year 1853, grouped by month
```{r Kolera_1853_mutated_words, warning=FALSE}
kolera_1853_tidy %>%
  mutate(m=month(timestamp)) %>%
  filter(str_detect(word, "\\bkolera[a-zæøå]+|\\bstolera[a-zæøå]+|\\bkolrra[a-zæøå+]"))%>%
    select(m,word)  %>%
  count(word, m)-> kolera_1853_mutated_words
```

### Creating the graph with points and lines
```{r graph, warning=FALSE}
kolera_1853_mutated_words %>%
  filter(str_detect(word, "\\bkolera[a-zæøå]+|\\bstolera[a-zæøå]+|\\bkolrra[a-zæøå+]")) %>%
  mutate(word = str_replace(word,"\\bkolera[a-zæøå,0-9,]+|\\bstolera[a-zæøå]+|\\bkolrra[a-zæøå]+[:graph:]*", "kolera")) %>%
    select(m, word) %>%
  count(word,m) %>%
  ggplot(aes(x = m, y = n, line = word, color = word)) +
  geom_line(color = "hotpink", size = 2) +
  geom_point(color = "hotpink", size = 5) +
  labs(
      title = "Occurence of any word in relation to cholera",
      subtitle = "Including OCR misreads", 
      x = "Month",
      y = "Number")+
  scale_x_continuous(breaks = 1:12)

ggsave("figures/graph.png")
```


