---
title: "Danish Kings"
author: "Frej Søgård Primdahl, Lucas Emil Fomsgaard Christensen, Stine Telling and Sidsel Skovhus Andersen"
date: "2025-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

Make sure to first create an  `.Rproj` workspace with a `data/` folder where you place either your own dataset or the provided `kings.csv` dataset. 

1. Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in plain text editor to see)

List what is the

separator: comma  

2. Create a `kings` object in R with the different functions below and inspect the different outputs.
-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=TRUE}
# FILL IN THE CODE BELOW and review the outputs
library(tidyverse)
kings1 <- read.csv("data/kings.csv")

kings2 <- read_csv("data/kings.csv", na="NULL")
head(kings2)
kings3 <- read.csv2("data/kings.csv")

kings4 <- read_csv2( "data/kings.csv")
```
Answer: 
1. Which of these functions is a `tidyverse` function? Read data with it below into a `kings` object: Answer: kings2 <- read_csv("data/kings.csv") is a tidyverse function, since it creates a tidy tibble.
2. What is the result of running `class()` on the `kings` object created with a tidyverse function: Answer: By running 'class()' on the 'kings2' object created with a tidyverse function, we can see that the values are in characters.
3. How many columns does the object have when created with these different functions?: Answer: By running 'ncol()' on the 'kings2' object created with a tidyverse function, we can see there is 11 columns.
4. Show the dataset so that we can see how R interprets each column: Answer: By using the head() function, R shows the first 6 rows of the dataset. It will show up below when running the head() function.

```{r eval = TRUE}
# COMPLETE THE BLANKS BELOW WITH YOUR CODE, then turn the 'eval' flag in this chunk to TRUE.

kings <- kings2

class("kings2")

ncol(kings2)

head(kings2)


```


## Calculate the duration of reign for all the kings in your table
You can calculate the duration of reign in years with `mutate` function by subtracting the equivalents of your `startReign` from `endReign` columns and writing the result to a new column called `duration`. But first you need to check a few things:

* Is your data messy? Fix it before re-importing to R
* Do your start and end of reign columns contain NAs? Choose the right strategy to deal with them: `na.omit()`, `na.rm=TRUE`, `!is.na()`


Create a new column called  `duration` in the kings dataset, utilizing the `mutate()` function from tidyverse. Check with your group to brainstorm the options.

```{r }
# YOUR CODE
na.omit("kings.csv")
kings_duration <- kings2 %>% 
  select(Name, Start_year, End_year) %>% 
  mutate(duration= End_year- Start_year, na.rm=TRUE)

kings_duration
```

## Calculate the average duration of reign for all rulers
Do you remember how to calculate an average on a vector object? If not, review the last two lessons and remember that a column is basically a vector. So you need to subset your `kings` dataset to the `duration` column. If you subset it as a vector you can calculate average on it with `mean()` base-R function. If you subset it as a tibble, you can calculate average on it with `summarize()` tidyverse function. Try both ways!

* You first need to know how to select the relevant `duration` column. What are your options?
* Is your selected `duration` column a tibble or a vector? The `mean()` function can only be run on a vector. The `summarize()` function works on a tibble.
* Are you getting an error that there are characters in your column? Coerce your data to numbers with `as.numeric()`.
* Remember to handle NAs: `mean(X, na.rm=TRUE)`


```{r}
# YOUR CODE
kings2 %>% 
  filter(!is.na(Start_year)) %>% 
  mutate(duration = End_year - Start_year) %>% 
  summarise((average=mean(duration, na.rm=TRUE)))
```

## How many and which kings enjoyed a longer-than-average duration of reign?

You have calculated the average duration above. Use it now to `filter()` the `duration` column in  `kings` dataset. Display the result and also  count the resulting rows with `count()`

```{r}
# YOUR CODE
filter(kings_duration,duration>19.57692)
count(filter(kings_duration,duration>19.57692))
```


## How many days did the three longest-ruling monarchs rule? 

* Sort kings by reign `duration` in the descending order. Select the three longest-ruling monarchs with the `slice()` function
* Use `mutate()` to create `Days` column where you calculate the total number of days they ruled
* BONUS: consider the transition year (with 366 days) in your calculation!

```{r}
# YOUR CODE
kings %>% 
mutate(duration = End_year- Start_year) %>% 
  arrange(desc(duration)) %>% 
  slice(1:3) %>% 
  mutate(days = duration * 365) %>% 
  mutate(LeapDays = duration * 366) %>% 
  select(Name,duration, days, LeapDays)
```


## Challenge: Plot the kings' duration of reign through time

What is the long-term trend in the duration of reign among Danish monarchs? How does it relate to the historical violence trends ? 

* Try to plot the duration of reign column in `ggplot` with `geom_point()` and  `geom_smooth()`
* In order to peg the duration (which is between 1-99) somewhere to the x axis with individual centuries, I recommend creating a new column `midyear` by adding to `startYear` the product of `endYear` minus the `startYear` divided by two (`startYear + (endYear-startYear)/2`).
* Now you can plot the kings dataset, plotting `midyear` along the x axis and `duration` along y axis
* BONUS: add a title, nice axis labels to the plot and make the theme B&W and font bigger to make it nice and legible!

```{r}
# YOUR CODE
glimpse(kings2)
kings <- kings2

kings %>% 
  mutate(duration = End_year - Start_year) %>% 
  mutate(Midyear = Start_year + duration/2) %>% 
  ggplot(aes(x = Midyear, y = duration))+
  geom_point()+
  geom_smooth()+
  theme_bw()
  
```

And to submit this rmarkdown, knit it into html. But first, clean up the code chunks, adjust the date, rename the author and change the `eval=FALSE` flag to `eval=TRUE` so your script actually generates an output.
Well done!


